---
title: "PK Dataset Analysis"
date: "`r Sys.Date()`"
output:
  officedown::rdocx_document:
    toc: true
    toc_depth: 3
    fig_width: 6.5
    tables:
      style: Table
      layout: autofit
      width: 1.0
      conditional:
        first_row: true
        first_column: false
        last_row: false
        last_column: false
        no_hband: true
        no_vband: false
params:
  data_path:
    label: "Dataset"
    value: "dataset.csv"
    input: file
  meta_data_path:
    label: "Metadata"
    value: "metadata.csv"
    input: file
  bins:
    label: "Bins for summary statistics"
    value: 11
    input: slider
    min: 1
    max: 20
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  comment = "#>",
  dev = "ragg_png",
  fig.align = "center",
  fig.width = 6.5
)
library(tidyverse)
library(patchwork)
library(nonmem.utils)

is_nonmem_file <- grepl(".tab", params$data_path)
if (is_nonmem_file) {
  data <- readr::read_table(params$data_path, na = c("NA", "N/A", "", "."), skip = 1)
} else {
  data <- readr::read_csv(params$data_path, na = c("NA", "N/A", "", "."))
}
meta_data <- readr::read_csv(params$meta_data_path, na = c("NA", "N/A", "", "."))
```
  
## Data inventory

__Table__: Subjects, doses and observations

```{r}
#| echo: false
#| results: asis
#| warning: false

inventory_list <- data_inventory(data, meta_data)
for (inventory_name in names(inventory_list)) {
  cat(inventory_name, "\n\n")
  cat(inventory_list[[inventory_name]] |> knitr::kable(), sep = "\n")
  cat("\n\n")
}
```


### Covariates

__Table__: Summary of continuous covariates

```{r}
#| echo: false
#| results: asis
#| warning: false

cov_list <- cov_inventory(data, meta_data)
for (cov_name in names(cov_list)) {
  cat(cov_name, "\n\n")
  cat(cov_list[[cov_name]] |> knitr::kable(), sep = "\n")
  cat("\n\n")
}
```

__Table__: Summary of categorical covariates

```{r}
#| echo: false
#| results: asis
#| warning: false

knitr::kable(cat_inventory(data, meta_data))
```

## Time Profiles

```{r}
#| echo: false
#| results: asis
#| warning: false

tp_plots <- time_profile(data, meta_data, bins = params$bins)

for (tp_name in names(tp_plots)) {
  cat(paste0("__Figure__: Time profile in linear scale for ", tp_name, "\n\n"))
  print(tp_plots[[tp_name]]$Linear)
  cat(paste0("__Figure__: Time profile in logarithmic scale for ", tp_name, "\n\n"))
  print(tp_plots[[tp_name]]$Log)
  cat(paste0("__Figure__: Time profile of BLQ proportions for ", tp_name, "\n\n"))
  print(tp_plots[[tp_name]]$`Percent BLQ`)
}
```

## Time After Dose Profiles

```{r}
#| echo: false
#| results: asis
#| warning: false

tp_plots <- tad_profile(data, meta_data, bins = params$bins)

for (tp_name in names(tp_plots)) {
  cat(paste0("__Figure__: Time after dose profile in linear scale for ", tp_name, "\n\n"))
  print(tp_plots[[tp_name]]$Linear)
  cat(paste0("__Figure__: Time after dose profile in logarithmic scale for ", tp_name, "\n\n"))
  print(tp_plots[[tp_name]]$Log)
  cat(paste0("__Figure__: Time after dose profile of BLQ proportions for ", tp_name, "\n\n"))
  print(tp_plots[[tp_name]]$`Percent BLQ`)
}
```


## Covariate distributions and correlations

__Figure__: Pairs plot of covariates

```{r}
#| echo: false
#| warning: false

cov_plot(data, meta_data)
```


__Table__: Spearman correlation and ANOVA regression coefficients estimated between covariates

```{r}
#| echo: false
#| results: asis
#| warning: false

cov_table <- cov_cor(data, meta_data)
knitr::kable(cov_table)
```

## Appendix

### Dictionary

```{r}
#| echo: false
#| results: asis
#| warning: false

knitr::kable(meta_data)
```

### Assessment of data ranges

```{r}
#| echo: false
#| results: asis
#| warning: false

knitr::kable(check_ranges(data, meta_data))
```

### Tables of time profiles summaries

