---
title: "Goodness of Fit Analysis"
date: "`r Sys.Date()`"
output:
  officedown::rdocx_document:
    toc: true
    toc_depth: 3
    fig_width: 6.5
    tables:
      style: Table
      layout: autofit
      width: 1.0
      conditional:
        first_row: true
        first_column: false
        last_row: false
        last_column: false
        no_hband: true
        no_vband: false
params:
  data_path:
    label: "Dataset"
    value: "dataset.tab"
    input: file
  meta_data_path:
    label: "Metadata"
    value: "metadata.csv"
    input: file
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  comment = "#>",
  dev = "ragg_png",
  fig.align = "center",
  fig.width = 6.5
)
library(tidyverse)
library(patchwork)
library(nonmem.utils)

is_nonmem_file <- grepl(".tab", params$data_path)
if (is_nonmem_file) {
  data <- readr::read_table(params$data_path, na = c("NA", "N/A", "", "."), skip = 1)
} else {
  data <- readr::read_csv(params$data_path, na = c("NA", "N/A", "", "."))
}
data$LLOQ <- 10
meta_data <- readr::read_csv(params$meta_data_path, na = c("NA", "N/A", "", "."))
```
  
## Observed vs predicted

__Figure__: Observed vs population and individual predictions in linear scale (left) and logarithmic scale (right)

```{r}
#| echo: false
#| results: asis
#| warning: false

dv_plot <- dv_preds(data, meta_data)
wrap_plots(
  dv_plot + guides(color = guide_legend(ncol = 2)),
  gg_log(dv_plot) + guides(color = "none"),
  ncol = 2
)
```


## Conditional Weighted Residuals

__Figure__: Histogram (left) and QQ-plot (right) of conditional weighted residuals (CWRES)

```{r}
#| echo: false
#| results: asis
#| warning: false

wrap_plots(
  residual_hist(y_type = "cwres", data, meta_data),
  residual_qq(y_type = "cwres", data, meta_data),
  ncol = 2
)
```

__Figure__: Conditional weighted residuals (CWRES) vs time, time after dose, population predictions, and individual predictions.

```{r}
#| echo: false
#| results: asis
#| warning: false

wrap_plots(
  residual_plot(x_type = "time", y_type = "cwres", data, meta_data),
  residual_plot(x_type = "tad", y_type = "cwres", data, meta_data),
  residual_plot(x_type = "pred", y_type = "cwres", data, meta_data),
  residual_plot(x_type = "ipred", y_type = "cwres", data, meta_data),
  ncol = 2
)
```


## Normalized Prediction Distribution Errors

__Figure__: Histogram (left) and QQ-plot (right) of normalized prediction distribution errors (NPDE)

```{r}
#| echo: false
#| results: asis
#| warning: false
#| eval: !expr '"NPDE" %in% names(data)'

wrap_plots(
  residual_hist(y_type = "npde", data, meta_data),
  residual_qq(y_type = "npde", data, meta_data),
  ncol = 2
)
```

__Figure__: Normalized prediction distribution errors (NPDE) vs time, time after dose, population predictions, and individual predictions.

```{r}
#| echo: false
#| results: asis
#| warning: false
#| eval: !expr '"NPDE" %in% names(data)'

wrap_plots(
  residual_plot(x_type = "time", y_type = "npde", data, meta_data),
  residual_plot(x_type = "tad", y_type = "npde", data, meta_data),
  residual_plot(x_type = "pred", y_type = "npde", data, meta_data),
  residual_plot(x_type = "ipred", y_type = "npde", data, meta_data),
  ncol = 2
)
```

## Individual Time Profiles

```{r}
#| echo: false
#| results: asis
#| warning: false

ind_time_profiles(data, meta_data)
```
