---
title: "VPC Analyses"
date: "`r Sys.Date()`"
output:
  officedown::rdocx_document:
    toc: true
    toc_depth: 3
    fig_width: 6.5
    tables:
      style: Table
      layout: autofit
      width: 1.0
      conditional:
        first_row: true
        first_column: false
        last_row: false
        last_column: false
        no_hband: true
        no_vband: false
params:
  data_path:
    label: "Dataset"
    value: "dataset.csv"
    input: file
  meta_data_path:
    label: "Metadata"
    value: "metadata.csv"
    input: file
  bins:
    label: "Bins for summary statistics"
    value: 11
    input: slider
    min: 1
    max: 20
  ci:
    label: "Confidence Interval"
    value: 80
    input: slider
    min: 51
    max: 99
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  comment = "#>",
  dev = "ragg_png",
  fig.align = "center",
  fig.width = 6.5
)
library(tidyverse)
library(nonmem.utils)

is_nonmem_file <- grepl(".tab", params$data_path)
if (is_nonmem_file) {
  sim_data_content <- readLines(params$data_path)
  # Remove header rows
  sim_data_file <- tempfile(fileext = ".tab")
  header_rows <- grepl(pattern = "ID", sim_data_content)
  data_header <- head(sim_data_content[header_rows], 1)
  sim_data_content <- sim_data_content[!header_rows]
  # Remove Table rows
  table_rows <- grepl(pattern = "TABLE", sim_data_content)
  sim_data_content <- sim_data_content[!table_rows]
  # Re-write temp file
  writeLines(c(data_header, sim_data_content), file = sim_data_file)
  data <- readr::read_table(sim_data_file) |> filter(MDV == 0)
} else {
  data <- readr::read_csv(params$data_path, na = c("NA", "N/A", "", "."))
}
meta_data <- readr::read_csv(params$meta_data_path, na = c("NA", "N/A", "", "."))

# Run VPC analysis
all_vpc_plots <- vpc_plots(
  data,
  meta_data,
  x = pull_name("tad", meta_data),
  bins = params$bins,
  ci = params$ci / 100
)
```

## Context

Visual Predictive Checks (VPCs) are graphical tools to compare the observed data with the simulated data from a population pharmacokinetic model. 
The VPCs display the median and confidence intervals (`r params$ci` %) of the simulated data, as well as the observed data.


## Regular Visual Predictive Checks

__Figure__: VPC in linear scale

```{r}
#| echo: false
#| warning: false

all_vpc_plots$vpc
```

__Figure__: VPC in log scale

```{r}
#| echo: false
#| warning: false

all_vpc_plots$vpc |> gg_log(x = FALSE)
```


## Visual Predictive Checks of BLQ distribution

__Figure__: VPC of BLQ distribution

```{r}
#| echo: false
#| warning: false

all_vpc_plots$blq
```

## Prediction-corrected Visual Predictive Checks

__Figure__: pc-VPC in linear scale

```{r}
#| echo: false
#| warning: false

all_vpc_plots$pc_vpc
```

__Figure__: pc-VPC in log scale

```{r}
#| echo: false
#| warning: false

all_vpc_plots$pc_vpc |> gg_log(x = FALSE)
```


## Visual Predictive Checks of Normalized Prediction Distribution Errors (NPDE)

__Figure__: VPC of NPDE distribution

```{r}
#| echo: false
#| warning: false

all_vpc_plots$npde
```
