---
title: "Goodness of Fit Analysis"
date: "`r Sys.Date()`"
output:
  officedown::rdocx_document:
    toc: true
    toc_depth: 3
    fig_width: 6.5
params:
  dataFile:
    label: "Nonmem Table"
    value: "dataset.tab"
    input: file
  metadataFile:
    label: "Metadata"
    value: "metadata.csv"
    input: file
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  comment = "#>",
  dev = "ragg_png",
  fig.align = "center",
  fig.width = 6.5
)
library(flextable)
library(tidyverse)
library(pmplots)
add_col <- function(data, col_name){
  data[[col_name]] <- data[[col_name]] %||% 0
  return(data)
}
if(!is.null(params$metadataFile)){
  meta_data <- read.csv(params$metadataFile)
  mdv_name <- ifelse(
    sum(meta_data$Type %in% "mdv")==1, 
    meta_data$Name[meta_data$Type %in% "mdv"],
    "MDV"
    )
  blq_name <- ifelse(
    sum(meta_data$Type %in% "blq")==1, 
    meta_data$Name[meta_data$Type %in% "blq"],
    "BLQ"
    )
  dv_row <- meta_data |> filter(Type %in% "dv")
  dv_name <- dv_row$Label
  if(!is.na(dv_row$Unit)){
    dv_name <- paste0(dv_name, " [", dv_row$Unit, "]")
  }
}else{
  mdv_name <- "MDV"
  blq_name <- "BLQ"
  dv_name <- "value"
}

data <- read.table(params$dataFile, header = TRUE, skip = 1) |>
  add_col(mdv_name) |>
  add_col(blq_name) |>
  filter(.data[[mdv_name]] == 0, .data[[blq_name]] <= 0)
```
  
# Observed vs Predicted

__Figure__: Observed vs Population (left) and Individual (right) Predictions in linear scale

```{r}
#| message: false
patchwork::wrap_plots(
  dv_preds(
    df = data,
    xname = dv_name,
    yname = dv_name
  ),
  ncol = 2
)
```

```{=openxml}
<w:br w:type="page"/>
```

__Figure__: Observed vs Population (left) and Individual (right) Predictions in logarithmic scale

```{r}
#| message: false
patchwork::wrap_plots(
  dv_preds(
    df = data,
    xname = dv_name,
    yname = dv_name,
    loglog = TRUE
  ),
  ncol = 2
)
```

```{=openxml}
<w:br w:type="page"/>
```

# Conditional Weighted Residuals

Conditional Weighted Residuals (CWRES) are a measure of the difference between the observed and predicted values, adjusted for the variability in the data.

__Figure__: Histogram (left) and QQ plot (right) of CWRES

```{r}
#| message: false
patchwork::wrap_plots(
  cwres_hist(df = data),
  cwres_q(df = data),
  ncol = 2
)
```

```{=openxml}
<w:br w:type="page"/>
```

__Figure__: CWRES vs Time (left), Time After Dose (right) and Population Predictions (bottom)

```{r}
#| message: false
patchwork::wrap_plots(
  cwres_time(df = data),
  cwres_tad(df = data),
  ncol = 2
)
```
```{r}
#| message: false
cwres_pred(df = data)
```

```{=openxml}
<w:br w:type="page"/>
```

# Normalized Prediction Distribution Errors

The Normalized Prediction Distribution Errors (NPDE) is a simulation-based analysis useful to investigate the accuracy of the model predictions, focusing more on comparing observations with their own simulated distributions than on capturing the concentration-time profile.

__Figure__: Histogram (left) and QQ plot (right) of NPDE

```{r}
#| message: false
patchwork::wrap_plots(
  cwres_hist(df = data),
  cwres_q(df = data),
  ncol = 2
)
```

```{=openxml}
<w:br w:type="page"/>
```

__Figure__: NPDE vs Time (left), Time After Dose (right) and Population Predictions (bottom)

```{r}
#| message: false
patchwork::wrap_plots(
  cwres_time(df = data),
  cwres_tad(df = data),
  ncol = 2
)
```
```{r}
#| message: false
cwres_pred(df = data)
```
