---
title: "Covariate Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Covariate Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(nonmem.utils)
```


## Preamble

Covariate analysis in PopPK modeling is a critical step used to identify and quantify factors that explain variability in PK parameters across individuals in a population.

To perform such analysis, Nonmem result file (e.g. `.tab` or `.par` file) must include `ETA` variable(s).

Such variables are defined in the meta data dictionary as `Type`: `eta` as illustrated below:

| Name | Type | Label | Unit |
|------|------|-------|------|
| ETA1 | eta  | &eta;CL |    |
| ETA2 | eta  | &eta;V |     |

## Interactive analysis

To analyse interactively Nonmem results you can use the shiny app below:

```r
nonmem.utils::run_shiny("covariate-analysis")
```

## Quarto Report

To create a docx report of the covariate analysis you can use the quarto based function below:

```r
nonmem.utils::report_covariate_analysis(
data_path = "my/dataset.tab", 
meta_data_path = "my/dictionary.csv"
)
```

## Available analyses

### Etas distribution and correlation

To analyze etas distributions and potential correlations, the function `eta_plot()` can be used.

```{r}
eta_plot(data_501, meta_data_501)
```

### Covariates vs etas

To analyze potential correlations between etas and covariates, 2 functions can be used: `eta_cov_plot()` and `eta_cor()`.

The function `eta_cov_plot()` allows graphical analyses of the covariate distribution and correlation:

```{r}
eta_cov_plot(data_501, meta_data_501)
```

The function `eta_cor()` provides a table with 

- Spearman correlation test results between continuous covariates and etas
- ANOVA test results between categorical covariates and etas

```{r}
knitr::kable(eta_cor(data_501, meta_data_501))
```
