---
title: "GOF Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GOF Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(nonmem.utils)
```

## Preamble

Goodness-of-fit (GOF) analysis is a key diagnostic tool used to evaluate how well a PopPK model describes the observed data. 
It helps determine whether the model is appropriate, reliable, and suitable for prediction

To perform such analysis, Nonmem result file (e.g. `.tab` file) must include usual Nonmem variables such as `PRED`, `IPRED`, `CWRES` and `NPDE`.

Such variables do not need to defined in the meta data dictionary as `PRED` and `IPRED` will use the `dv` type for labelling.


## Interactive analysis

To analyse interactively Nonmem results you can use the shiny app below:

```
nonmem.utils::run_shiny("gof-analysis")
```

## Quarto Report

To create a docx report of the GOF analysis you can use the quarto based function below:

```
nonmem.utils::report_gof_analysis(
data_path = "my/dataset.tab", 
meta_data_path = "my/dictionary.csv"
)
```

## Available analyses

### Observed vs predicted

The function `dv_preds()` plots observed vs individual and population predictions in both linear and log scales.

```{r}
dv_preds(data_501, meta_data_501)
```

To obtain the plot in log-log scale, the function `gg_log()` can be used as illustrated below:

```{r}
dv_preds(data_501, meta_data_501) |> 
  gg_log()
```

### Residuals

To assess residual distributions, 3 functions are available: `residual_hist()`, `residual_qq()` and `residual_plot()`.

- The function `residual_hist()` plots histogram of residuals

```{r}
residual_hist(y_type = "cwres", data = data_501, meta_data = meta_data_501)
```

- The function `residual_qq()` perform QQ-plots of residuals

```{r}
residual_qq(y_type = "cwres", data = data_501, meta_data = meta_data_501)
```

- The function `residual_plot()` plots residuals vs another variable to analyze potential trends

```{r}
patchwork::wrap_plots(
  residual_plot(x_type = "time", y_type = "cwres", data = data_501, meta_data = meta_data_501),
  residual_plot(x_type = "pred", y_type = "cwres", data = data_501, meta_data = meta_data_501)
)
```

### Individual time profiles

To review individual time profiles, the functions `ind_time_profiles()` and `ind_tad_profiles()` can be used as illustrated below. The functions leverage simple 1-, 2-, and 3-compartment models using the `mrgsolve` package and PK parameters Nonmem naming convention (e.g. if dataset includes only `CL` and `V`, the code assumes a 1-compartment model).

```{r}
ind_time_profiles(
  data = data_501 |> dplyr::filter(ID <= 6), 
  meta_data = meta_data_501
  )
```

```{r}
ind_tad_profiles(
  data = data_501 |> dplyr::filter(ID <= 6), 
  meta_data = meta_data_501
  )
```

