---
title: "Goodness of Fit Analysis"
date: "`r Sys.Date()`"
output:
  officedown::rdocx_document:
    toc: true
    toc_depth: 3
    fig_width: 6.5
params:
  dataFile:
    label: "Nonmem Table"
    value: "dataset.tab"
    input: file
  metadataFile:
    label: "Metadata"
    value: "metadata.csv"
    input: file
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  comment = "#>",
  dev = "ragg_png",
  fig.align = "center",
  fig.width = 6.5
)
meta_data <- read.csv(params$metadataFile)
dv_name <- "FIX activity [IU/L]"
data <- read.table(params$dataFile, header = TRUE, skip = 1) |>
  filter(MDV == 0)
library(flextable)
library(tidyverse)
library(pmplots)
```
  
# Observed vs Predicted

__Figure__: Observed vs Population (left) and Individual (right) Predictions in linear scale

```{r}
#| message: false
patchwork::wrap_plots(
  dv_preds(
    df = data,
    xname = dv_name,
    yname = dv_name
  ),
  ncol = 2
)
```

```{=openxml}
<w:br w:type="page"/>
```

__Figure__: Observed vs Population (left) and Individual (right) Predictions in logarithmic scale

```{r}
#| message: false
patchwork::wrap_plots(
  dv_preds(
    df = data,
    xname = dv_name,
    yname = dv_name,
    loglog = TRUE
  ),
  ncol = 2
)
```

```{=openxml}
<w:br w:type="page"/>
```

# Conditional Weighted Residuals

Conditional Weighted Residuals (CWRES) are a measure of the difference between the observed and predicted values, adjusted for the variability in the data.

__Figure__: Histogram (left) and QQ plot (right) of CWRES

```{r}
#| message: false
patchwork::wrap_plots(
  cwres_hist(df = data),
  cwres_q(df = data),
  ncol = 2
)
```

```{=openxml}
<w:br w:type="page"/>
```

__Figure__: CWRES vs Time (left), Time After Dose (right) and Population Predictions (bottom)

```{r}
#| message: false
patchwork::wrap_plots(
  cwres_time(df = data),
  cwres_tad(df = data),
  ncol = 2
)
```
```{r}
#| message: false
cwres_pred(df = data)
```

```{=openxml}
<w:br w:type="page"/>
```

# Normalized Prediction Distribution Errors

The Normalized Prediction Distribution Errors (NPDE) is a simulation-based analysis useful to investigate the accuracy of the model predictions, focusing more on comparing observations with their own simulated distributions than on capturing the concentration-time profile.

__Figure__: Histogram (left) and QQ plot (right) of NPDE

```{r}
#| message: false
patchwork::wrap_plots(
  cwres_hist(df = data),
  cwres_q(df = data),
  ncol = 2
)
```

```{=openxml}
<w:br w:type="page"/>
```

__Figure__: NPDE vs Time (left), Time After Dose (right) and Population Predictions (bottom)

```{r}
#| message: false
patchwork::wrap_plots(
  cwres_time(df = data),
  cwres_tad(df = data),
  ncol = 2
)
```
```{r}
#| message: false
cwres_pred(df = data)
```
